---
title: "Project - University Enrollment"
author: "ERICK@Guru"
date: '2023'
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practical Exam - University Enrollment

**Instructions**

● Use Python or R to perform the tasks required.

● Write your solutions in the workspace provided from your certification page.

● Include all of the visualizations you create to complete the tasks.

Visualizations must be visible in the published version of the workspace. Links to external visualizations will not be accepted.

● You do not need to include code unless the question says you must.

● You must pass all criteria to pass this exam.

**The full criteria can be found here.**

Background: You are working as a data scientist at a local University. The university started offering online courses to reach a wider range of students. The university wants you to help them understand enrollment trends. They would like you to identify what contributes to higher enrollment. **In particular, whether the course type (online or classroom) is a factor.** 1

```{r warning=FALSE, message=FALSE}
#Import libraries
library(tidyverse)
library(dplyr)
library(corrplot)
library(VIM)
library(cowplot)
library(parsnip)
library(ggplot2)
library(janitor)
library(rsample)
library(e1071)#svm model
library(DataExplorer)
library(report)
library(lmtest)#test the model#harvtest(model) funtion we want p-value>0.05
# shapiro.test(residuals(model)) we want p-value>0.05
theme_set(theme_test())
library(tinytex)
library(bookdown)
library(rmarkdown)
```

```{r  warning=FALSE, message=FALSE}
#Import dataset
enrol <-read_csv("university_enrollment_2306.csv")
head(enrol)
```

```{r warning=FALSE, message=FALSE}
glimpse(enrol)
str(enrol)
```

```{r warning=F, message=FALSE}
#check the number of missing values per column
as.data.frame(
 colSums(is.na(enrol))
)
```

### Data Manipulation & Cleaning

```{r}
#####Column 1
#no duplicates
sum(duplicated(enrol$course_id))

unique(enrol$course_type)#"classroom" "online"  #no missing detected
enrol$course_type <- as.factor(enrol$course_type)
#convert to a factor variable

#Enrollment_Count-Nothing to be changed

#nothing to be changed
unique(enrol$year)
#2018 2020 2016 2013 2019 2014 2021 2022 2011 2015 2012 2017
#Nothing to be amended

###Pre_score should be numeric###
enrol$pre_score <- as.numeric(enrol$pre_score)
sum(is.na(enrol$pre_score))
enrol$pre_score <- ifelse(is.na(enrol$pre_score), 0, enrol$pre_score)

###Post_Score should be numeric
class(enrol$post_score)
sum(is.na(enrol$post_score))
enrol$post_score <- ifelse(is.na(enrol$post_score), 0, enrol$post_score)
##########
unique(enrol$pre_requirement)#Replace missing with NONE
enrol$pre_requirement<-ifelse(is.na(enrol$pre_requirement),
                              "None",enrol$pre_requirement)
# enrol$pre_requirement <- as.factor(enrol$pre_requirement)
# class(enrol$pre_requirement)
# levels(enrol$pre_requirement)

################
unique(enrol$department)#Replace Math-Mathemathics#no missing values
enrol$department <- ifelse(enrol$department=="Math",
                           "Mathematics",
                           enrol$department)
# enrol$department <- as.factor(enrol$department)
```

```{r}
#To this far we have corrected our variables as per the instruction
glimpse(enrol)
```

# Tasks: Our Objectives

**Write your answers in your workspace.**

## 1. For every column in the data:

a.  State whether the values match the description given in the table above.

```{r}
str(enrol)
#All the variables has been corrected as per the description

```

b\. State the number of missing values in the column.

```{r}
#Number of missing values per column
colSums(is.na(enrol))

```

c\. Describe what you did to make values match the description if they did not match.

-   course_id- the column was OK

-   course_type- Nothing to be corrected

-   year-Nothing to be corrected

-   enrollment_count-Nothing to be changed

-   pre_score- I changed the variable to numeric, then I replace the Missing values with "0" Zero

-   post_score-I replace the Missing values with "0"

-   pre_requirement- I Replace missing with "`NONE`"

-   department- There were no missing values, I Replace '*`Math`*'- with "*`Mathemathics`*"

## 2. Describe the distribution of the enrollment counts. Your answer must include a visualization that shows the distribution.

```{r}
d1 =enrol %>% 
  group_by(year) %>% 
  summarise(Total_yearly=sum(enrollment_count))

range(d1$Total_yearly)
hist(d1$Total_yearly, col='orange')

p1 <- d1 %>% 
  ggplot(aes(x=factor(year), y=Total_yearly))+
  geom_col(position = 'dodge', fill=rainbow(12)) +
  scale_y_continuous(labels = scales::comma)+
  geom_text(aes(label=Total_yearly),
            position = position_stack(0.4),
            vjust=-.2,
            hjust=-.04
            )+
  ggtitle('Distribution of Enrollment Per Year')+
  coord_flip()
p1
```

The **Range** of the Total enrollment per year is `32696 - 38897,` and the histogram of the Total enrollment shows that the Enrollment Yearly is normally distributed.

```{r}
d <- enrol %>% 
     group_by(course_type) %>% 
    summarise(Total_enrol= sum(enrollment_count)) %>% 
    arrange(desc(Total_enrol))
# DT::datatable(d)

p0 <- d %>% 
  ggplot(aes(x=course_type, y=Total_enrol))+
  geom_col(fill=rainbow(2)) +
  scale_y_continuous(labels = scales::comma)+
  geom_text(aes(label=Total_enrol),
            vjust=4,
            hjust=-.3
            ) +
  ggtitle('Distribution of Enrollment Per Course Type')



d2 <- enrol %>% 
  group_by(pre_requirement) %>% 
  summarise(Total_P=sum(enrollment_count))
  
p2 <- d2 %>% 
  ggplot(aes(x=factor(pre_requirement), y=Total_P))+
  geom_col(fill=rainbow(3)) +
  geom_text(aes(label=Total_P),
            position = position_stack(0.5)
                        )+
  ggtitle('Distribution of Enrollment Per Requirement')

d3 <- enrol %>% 
  group_by(department) %>% 
  summarise(Total_D=sum(enrollment_count))
p3 <-  d3 %>%  
   ggplot(aes(x=department, y= Total_D, fill=department))+
   geom_col()+
   geom_text(
     aes(label=Total_D),
     position = position_stack(0.6)
   )+
   theme(legend.position = 'none')+
   ggtitle('Distribution of Enrollment  Count Per Department')

plot_grid(p0, p2, p3, labels = "auto")
```

**In the above charts.**

-Most students prefer `Online` Classes

-When the course requirement is `NONE`, the enrollment will also goes up.

-Most students prefer `Technology` courses

```{r}
enrol %>%
select(starts_with(c('enrol', 'year', 'pre_s', 'post'))) %>%
cor() %>%
corrplot(type = "upper", addCoef.col = "black", diag=FALSE)
```

We have the correlation coefficients in each box. Positive correlations are in blue. Negative correlations are in red.

r \>= 0.5 large or strong association

r = 0.3 medium association

r = 0.1 small or weak association

**Summary of correlations:**

In respect to the above conditions, we can conclude that , the year of enrollment, the pre-score and post-score value doesn't affect the enrollment outcome.

## 3. Create a visualization that shows how many courses were of each type. Use the visualization to:

```{r}
plot_grid(p0, p2, p3, labels = "auto")
```

a\. State which type of course has the most observations

`-Online Courses attract more number of learners`

`-Most learners prefer Technology Courses`

`-When their are NO, or NONE requirement is needed, most learners enroll.`

b\. Explain whether the observations are balanced across the types.

`-The observations are not balanced at all, for instance Online enrollment is 347442 while Classroom enrollment is 80954.`

## 4. Describe the relationship between `course type` and the `enrollment count`. Your answer must include a visualization to demonstrate the relationship.

```{r}
d %>% 
  ggplot(aes(x=course_type, y=Total_enrol))+
  geom_col(fill=rainbow(2)) +
  scale_y_continuous(labels = scales::comma)+
  geom_text(aes(label=Total_enrol),
            vjust=4,
            hjust=-.3
            ) +
  ggtitle('Distribution of Enrollment Per Course Type')

enrol %>% 
  ggplot(aes(course_type,enrollment_count)) +
  geom_boxplot()
```

`-Majority of learners prefer Online compared to Classroom ,`

## 5. The university wants to predict how many students will enroll in a course. State the type of machine learning problem that this is (regression/ classification/ clustering).

`-This is a Regression Problem`

# `Model Development and fitting`

## 6. Fit a baseline model to predict how many students will enroll using the data provided. You must include your code.

```{r}
names(enrol)
```

```{r}
#
set.seed(13579)
#split train/test set
split <- initial_split(enrol, prop = 0.75)
train <- training(split)
test <- testing(split)
```

```{r}
#Base line Model
#Linear Regression Model
enrol$department1 <- NULL
enrol$course_id <- NULL
model <- lm(formula = enrollment_count~ scale(year)+scale(pre_score) +
              scale(post_score)+
              factor(course_type) + factor(pre_requirement)+
              factor(department),  data=train)
summary(model)
report(model)
pred <- predict(model, test) %>%
  round(0) 

data.frame(Actual=test$enrollment_count,Prediction= pred)
caret::RMSE(pred,test$enrollment_count)
caret::R2(pred,test$enrollment_count)
```

`The generic interpretation for each of our coefficients is for every one unit increase in the x variable, the y variable (enrollment_count) increases by beta units.`

What can we derive from our correlation analysis and how can we use this to inform **`enrollment_count`**?

We will just focus on the relationship between **`enrollment_count`**and each independent (x) variable. **What's not working in `enrollment`?**

1\. `Year— significant.`

`2. Department— significant`

`3. factor(pre_requirement)Intermediate — significant.`

**What is working with marketing?**

`1.Course Tpye`

`2.Pre Requirement  NONE`

**What is not as impactful?**

`1. Pre score`

`2.Post score`

## 7. Fit a comparison model to predict how many students will enroll using the data provided. You must include your code.

```{r}
#comparison Model
#Support Vector Regression Model

model2 <- svm(formula = enrollment_count~year+pre_score +post_score+
              factor(course_type) + factor(pre_requirement)+
              factor(department),  data=train)
# summary(model2)
pred2 <- predict(model2, test) %>% round(0)

data.frame(test$enrollment_count, pred2)

caret::RMSE(pred2,test$enrollment_count)
caret::R2(pred2,test$enrollment_count)
```

## 8. Explain why you chose the two models used in parts 6 and 7.

`I choose the Two models , so that I check the one with more accurate resulst`

## 9. Compare the performance of the two models used in parts 6 and 7, using any method suitable. You must include your code.

R-Square adjusted for Linear Regression Model is `0.9999398`, compared to SVM Model with R2 of `0.9962162` , with the R2 , we can conclude that the **Linear Model** Performed better than **SVM**.

## 10. Explain which model performs better and why.

**Linear Model** Performed better than **SVM. The RMSE of Linear Model is** `0.2975784` **smaller than for SVM Model at** `2.334824`
